---
title: "Lab 6"
author: "Caroline Shepherd"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test).

HO: Equal ranks (or medians)
(Null- there is no separation by rank in these populations)

HA: Ranks are not equal (or medians are not equal)


### Create samples

'set.seed()' to create "pseudo random, but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributions using the 'hist() function'

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the argument 'paired = TRUE' within the t.test() function.
```

This P-value means that IF our samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is `r round(ttest$p.value,2)` probability of finding sample means at least as different as mine by random chance

A more comprehensive explanation of differences between groups.... actually means data, data spread, effect size.... there is no significant difference (t(df) = t-statistic value, p = ).

### Mann Whitney U rank based test

We use the 'wilcox.test()' function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that if our samples are drawn from a population with the same median, there is a 28% chance of finding sample medians at least as different as ours by random chance.

No significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p value = `r round(mwu$p.value, 2)`).

The medians of group A (M = 22.4 g) and group B (M = 31.2 g) did not differ significantly (Mann Whitney U,...)


## Simple linear regression

Two v*continous quantitative* ariables that you are trying to understand the relationsihp between and you think that a linear relationship makes sense.

Use the 'penguins" data from {palmerpenguins}, exploring flipper length (x) versus body mass (y).

```{r}
# Use view(penguins) to look at it 
# Make an exploratory scatterplot of penguin flipper length (x-axis) versus body mass (y-axis)

view(penguins)

ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")
```
### Find the model estimates (coefficients)

Use the 'lm()' function fo rlinear model.
body_mass ~ <<<< tilde (not a minus sign)

```{r}
penguins_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

```

Model coefficients: `r penguins_lm$coefficients`
Flipper length coefficient: `r round(penguins_lm$coefficients[2], 2)`

### Use 'broom::tidy()' to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguins_lm)

penguin_lm_tidy

```

### Use broom::glance() to get output of model fit and statistical significance

```{r}
penguin_lm_glance <- glance(penguins_lm)

penguin_lm_glance
```
### Use equatiomatic::extract_eq() to return the model in LaTeX

```{r}
extract_eq(model = penguins_lm, use_coefs = TRUE)
```

### Use plot() function to automatically produce diagnostic plots

```{r}
plot(penguins_lm)
```


